---
title: "ADA_Homework2"
author: "Caitlin O'Connell"
date: "9/5/2022"
output: html_document
---

## Installing packages
```{r}
# install.packages("readr")
# install.packages("dplyr")
# install.packages("openxlsx")

# for reading csv file (read_csv())
library(readr)

# for data management
library(dplyr)

# for exporting excel files
library(openxlsx)

```

## Number 1
Import class one survey data from our Github site (https://github.com/kijohnson/ADA) 
(The dataset is called ‘Class 1 Survey Fall 2022.csv’), calling the R dataframe that you create C1survey.
```{r}
C1survey <- read_csv(
  "https://raw.githubusercontent.com/kijohnson/ADA/master/Class%201%20Survey%20Fall%202022.csv")

```

## Number 2
Determine: a. the number of observations (i.e. the number of people who filled out the survey) and 
b. the number of variables in the dataframe.
```{r}
# finding the dimensions of the data set
dim(C1survey)
```

a. 42 people filled out the survey (42 observations)
b. 27 variables in the dataframe

## Number 3
a. Rename the column variables to something shorter and that is descriptive of what the variable is about 
(for example like_dogs for the ‘Do you like dogs?’ question variable) and 
b. Display that they are renamed.
```{r}
# renaming the column variables using rename function
C1survey <- C1survey %>%
  rename("like_cats" = "Do you like cats?") %>%
  rename("like_dogs" = "Do you like dogs?") %>%
  rename("item_desert_island" = 
           "If you were stranded on a desert island, what is the one item you want to have with you?") %>%
  rename("favorite_day" = "What is your favorite day of the week?") %>%
  rename("lark_owl_humbird" = "Are you a lark, an owl, or a hummingbird?") %>%
  rename("favorite_food" = "What is your favorite food?") %>%
  rename("favorite_drink" = "What is your favorite drink?") %>%
  rename("favorite_season" = "What is your favorite season?") %>%
  rename("favorite_Month" = "What is your favorite Month?") %>%
  rename("hobby" = "What is one hobby that you enjoy?") %>%
  rename("degree_program" = "What is your degree program?") %>%
  rename("specialization" = "What is your primary specialization?") %>%
  rename("stats_software" = "What is your primary statistical software?") %>%
  rename("coding_comfort" = "How comfortable are you with coding?") %>%
  rename("favorite_number" = "What is your favorite number?") %>%
  rename("home_country" = "What is your home country?") %>%
  rename("home_state" = "What is your home state?") %>%
  rename("home_city" = "What is your home city?") %>%
  rename("education_level" = "What is your highest education level?")

# renaming longer names using the names function
names(C1survey)[5] <- c("life_slogan")
names(C1survey)[16] <- c("R_experience")
names(C1survey)[18:20] <- c("coding_time","learning_goals","interest_area")
names(C1survey)[22:23] <- c("birth_day","birth_month")

# displaying they are renamed
View(C1survey)
```

## Number 4
Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.

```{r}
# determining the type of variables in the dataset
sapply(C1survey,class)
```

```{r}
# defining values to display the number of character, numerical, and factor 
num_character <- 23
num_numerical <- 4
num_factor <- 0
```

## Number 5
a. Using code, check the bday and bmonth variables for any unusual or missing values. 
If you find any, 
b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. 
*Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}
# converting the bday and month variables to numeric
C1survey$birth_day <- as.numeric(C1survey$birth_day)
C1survey$birth_month <- as.numeric(C1survey$birth_month)
```

```{r}
# summarizing the two variables (part a)
summary(C1survey$birth_day)
summary(C1survey$birth_month)
```

```{r}
# I am going to keep the NA's and adding na.rm=TRUE to my code finding the median.

# finding the median birth day and month
median_bday <- median(C1survey$birth_day, na.rm = TRUE)
median_bmonth <- median(C1survey$birth_month, na.rm = TRUE)

# displaying medians
median_bday
median_bmonth
```

## Number 6
a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter=12/1 to 2/29, spring=3/1 to 5/31, summer=6/1 to 8/31, fall=9/1 to 11/30). 

b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct 

c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.
```{r}
# creating a variable called bseason 
# gives the season in which respondents were born

C1survey<-C1survey %>%
 mutate(bseason=if_else(birth_month==1| birth_month == 2, "winter", 
                             if_else(birth_month == 3|birth_month == 4| birth_month == 5, "spring",
                                     if_else(birth_month == 6|birth_month == 7| birth_month == 8, "summer",
                                             if_else(birth_month == 9|birth_month==10|birth_month==11, "fall",
                                                     if_else(birth_month == 12, 
                                                             "winter", NULL)))))) %>%
  mutate(bseason = as.factor(bseason))
summary(C1survey$bseason)
```

```{r}
# table of Seasons in columns and birth month in row
season_table <- table(C1survey$birth_month, C1survey$bseason, useNA = "always")
```

```{r}
# Summing the columns to answer the question of how many classmates were born in each season
addmargins(season_table, margin = seq_along(dim(season_table)), FUN = sum, quiet = TRUE)
```

## Number 7
Pick your favorite variable to analyze, come up with a question you want to answer with that variable, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).

```{r}

```

